Toward Supervised Reinforcement Learning with Partial States for Social HRI

### 2017-08-23 paper.tex
Further updates

Initial knowledge is often a crucial step to allow an algorithm to learn an
efficient action policy. This knowledge, originating from human expertise, can
be exploited in many ways.

@@

Approaches from animal conditioning have also been transfered to Machine
Learning. For example, the TAMER framework~\cite{knox2009interactively} is an
adaptation of ``shaping'', tuning an animal's behaviour by providing rewards to
guide it to the desired behaviour. Unlike other RL setups, with TAMER, the
reward is not provided by the environment but by a human observer. TAMER derives
a reward function from the human feedbacks and uses this function to evaluate
the current behaviour. In a similar approach, MacGlashan et al. proposed COACH
in~\cite{macglashan2017interactive}. COACH assumes that human rewards represent
the advantage function, i.e. how much the current action is better than the
current policy. With COACH, the teacher can adapt the rewards to the current
policy and performance of the agent, for example starting to positively reward
suboptimal behaviours that are better than the current policy and later, as the
agent improves its behaviour, reward these same suboptimal behaviours negatively
and only the optimal ones positively.

@@

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
learning based on human feedback as illustrated in Figure~\ref{fig:comparison}.
In Autonomous Learning, the agent has first to discover its environment to start
gathering relevant feedbacks, leading to a low initial performance and a slow
improvement in early stage of learning. With a teaching based on human
feedbacks, the teacher can quickly provide feedbacks on actions to reach more
quickly an efficient policy. However, as the teacher only provide feedbacks and
cannot prevent the agent to make mistakes, the initial performance can be poor.
On the other hand, SPARC, with the control of the teacher over the executed
actions can prevent the agent to make mistakes in early stages, achieving a high
performance even at the start of the learning. The total blending between
autonomous execution of actions and demonstrations and the control over
\emph{every} actions executed by the agent are the specificities of SPARC
compared to other approaches using humans to guide the learning of an agent.

@@

In this paper we propose to use a human user to highlight the relevant features
of the environment to reduce the state dimension of the points only to relevant
information.  We introduce the concept of a partial state, a sliced version of
the state defined only on a subset of the dimensions of the state. This shifts
the (state, action) pair paradigm to (partial state, action)  and in the case of
RL, the tuple (state, action, reward) to (partial state, action, reward). This
allows to compare the current state and the datapoints only on relevant features
for action selection.

@@

The normalisation of each dimension of the state allows distances to be
comparable as values on all dimensions have the same range. Additionally taking
the average quadratic distance over each defined dimension allows to compare
distances even when states are defined on a different number of dimensions.

%When selecting an action, the agent can also use
%the dimensions of the close state $s'$ used for the selection to indicate the
%supervisor which parts of the state have been used for the action selection.

>>>>

Initial knowledge is often crucial to allow an algorithm to learn an
efficient action policy. This knowledge, originating from human expertise, can
be exploited in many ways.

@@

Agents learning in environments providing rewards can also benefit from
additional rewards from human teachers. Depending on the task and the
environment, different ways exist to combine rewards from multiple sources, and
studies show that augmenting rewards from the environment by human ones can
speed up the learning and reduce the number of undesired behaviours~
\cite{griffith2013policy,knox2010combining,judah2010reinforcement}.

When environments do not supply rewards, they can be replaced by human ones. For
example, the TAMER framework~\cite{knox2009interactively} derives a reward
function from the human feedbacks and uses this function to evaluate the current
behaviour. In a similar approach, MacGlashan et al. proposed COACH
in~\cite{macglashan2017interactive}. COACH assumes that human rewards represent
the advantage function, i.e. how much the current action is better than the
current policy, allowing to adapt the reward function to the strategies used and
the current state of the learner.

@@

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
learning based on human feedback as illustrated in Figure~\ref{fig:comparison}.
In Autonomous Learning, the agent has first to discover its environment to start
gathering relevant feedbacks, leading to a low initial performance and a slow
improvement in early stage of learning. With a teaching based on human
feedbacks, the teacher can quickly provide feedbacks on actions to reach more
quickly an efficient policy. However, as the teacher only provide feedbacks and
cannot prevent the agent to make mistakes, the initial performance can be poor.
On the other hand, SPARC, with the control of the teacher over the executed
actions can prevent the agent to make mistakes in early stages, achieving a high
performance even at the start of the learning.

Compared to similar algorithms
\cite{chernova2009interactive,walsh2010generalizing}, which allows the agent to
request demonstrations and/or the teacher to provide demonstrations to correct
mistakes made by the robot, SPARC allows the teacher to correct any  action
before its execution, thus reducing importantly the risk of the agent making
errors. The total blending between autonomous execution of actions and
demonstrations and the control over \emph{every} actions executed by the agent
are the specificities of SPARC.

@@

In this paper we propose to use a human user to highlight the relevant features
of the environment to reduce the state dimension of the points only to relevant
information. We introduce the concept of a partial state, a sliced version of
the state defined only on a subset of the dimensions of the state. This shifts
the (state, action) pair paradigm to (partial state, action)  and in the case of
RL, the tuple (state, action, reward) to (partial state, action, reward). This
allows to compare the current state and the datapoints only on relevant features
for action selection. With this instance-based method~\cite{aha1991instance},
the algorithm can have a state abstraction allowing it to generalise even with a
few datapoints instead of the large numbers normally required to abstract
features from examples.

@@

The normalisation of each dimension of the state allows distances to be
comparable as values on all dimensions have the same range. Additionally taking
the average quadratic distance over each defined dimension allows to compare
similarities even when states are defined on a different number of dimensions.
